cardinalR: Generating interesting high-dimensional data structures

### 2025-05-04 paper-cardinalR.Rmd
Added gen_circle definition

>>>>

The function `gen_circle(n, p)` generates a $p$-dimensional dataset of $n$ observations, where the first two dimensions form a unit circle, and the remaining dimensions are structured sinusoidal extensions of the angular parameter with progressively smaller scale.

A latent angle variable $\theta$ is uniformly sampled from the interval $[0, 2\pi]$. Coordinates in the first two dimensions represent a perfect circle on the plane:

$$
X_1 = \cos(\theta), \quad X_2 = \sin(\theta)
$$

For dimensions $X_3$ through $X_p$, sinusoidal transformations of the angle $\theta$ are introduced. The first component is a scaling factor that decreases with the dimension index, defined as $\text{scale}_j = \sqrt{(0.5)^{j-2}}$ for $j = 3, \dots, p$. The second component is a phase shift that is proportional to the dimension index, specifically designed to decorrelate the curves, given by the formula $\phi_j = (j - 2) \cdot \frac{\pi}{2p}$.

Each additional dimension is computed as:

$$
X_j = \text{scale}_{j} \cdot \sin(\theta + \phi_j), \quad j = 3, \dots, p
$$

This structure retains a one-dimensional manifold (the circle) but allows it to be embedded non-linearly and decorrelatively in higher dimensions. The decreasing scale ensures that higher dimensions add less variation, preserving the dominance of the circular shape while adding complexity.
